---
title: "Data Preprocessing Housing Gradient Italy"
output: html_notebook
---

A dataset with information about housing price gradients and immigrant population in regional capitals in Italy. Data source:

Antoniucci, Valentina, and Giuliano Marella. "Housing price gradient and immigrant population: Data from the Italian real estate market." Data in brief 16 (2018): 794-798. https://doi.org/10.1016/j.dib.2017.12.018

Paper saved as: 1-s2.0-S2352340917307138-main.pdf

In this notebook, the data are pre-processed for packaging in `R`.

Load libraries:
```{r}
library(labelled)
library(lubridate)
library(readxl)
library(tidyverse)
```

Data saved as: 1-s2.0-S2352340917307138-mmc2.xlsx

Read data file (original is in Stata format):
```{r}
price_gradient <- read_xlsx("1-s2.0-S2352340917307138-mmc2.xlsx")
```

Get the column names and labels; these will be used in the documentation:
```{r}
data_names <- data.frame(names = c("capital_name",
                                   "grad_existing",
                                   "grad_new",
                                   "population",
                                   "log_population",
                                   "immigrants",
                                   "log_immigrants",
                                   "emp_retail_tourism",
                                   "emp_service",
                                   "female_emp_rate",
                                   "emp_rate",
                                   "income_per_capita",
                                   "log_income_per_capita",
                                   "density",
                                   "log_density",
                                   "transit_per_capita",
                                   "distance",
                                   "housing_surface",
                                   "mean_alt",
                                   "housing_units",
                                   "building_density"),
                         label = colnames(price_gradient))
```

Write description of variables in roxygen-style:
```{r}
data_names <- data_names %>% mutate(description = glue::glue("#'   \\item{{{names}}}{{{label}}}"))
```

Rename columns, which will change after converting data to factors:
```{r}
colnames(price_gradient) <- data_names$names
```

Preprocess data:
```{r}
price_gradient <- price_gradient %>%
  mutate(capital_name = factor(capital_name))
```

Delete unzipped excel file:
```{r}
unlink("1-s2.0-S2352340917307138-mmc2.xlsx")
```

Save to file:
```{r}
usethis::use_data(price_gradient,
                  overwrite = TRUE)
```
