---
title: "Industrial Readiness Index"
output: html_notebook
---

A dataset with information about distressed real estate market auctions in Italy. Data source:

Canesi, Rubina, and Giuliano Marella. "Data from RE distressed market: Properties auctions in Italy." Data in brief 18 (2018): 319-324. https://doi.org/10.1016/j.dib.2018.03.009

Paper saved as: 1-s2.0-S2352340918302105-main.pdf

In this notebook, the data are pre-processed for packaging in `R`.

Load libraries:
```{r}
library(labelled)
library(lubridate)
library(readxl)
library(tidyverse)
```

Data saved as: 1-s2.0-S2352340918302105-mmc2.xlsx

## Original data

Read data file:
```{r}
italy_auctions <- read_excel("1-s2.0-S2352340918302105-mmc2.xlsx", 
                             col_types = c("numeric", "numeric", "numeric", 
                                           "numeric", "numeric", "numeric", 
                                           "numeric", "numeric", "numeric", 
                                           "numeric", "numeric", "numeric", 
                                           "numeric", "numeric", "date"),
                             skip = 1)
```

Fix one date that was incorrectly coded:
```{r}
italy_auctions$DV[44] <- "2015-06-01 UTC"
```

Preprocess data frame:
```{r}
italy_auctions <- italy_auctions %>%
  rename(type_class = TC,
         gross_building_area = GBA,
         quality = Q,
         state_maintenance = SM,
         occupancy = Oc,
         location = Loc,
         income = Inc,
         delta_ntn = DNTN,
         re_activity_index = REAI,
         population = Pop,
         days_on_market = DD,
         number_auctions = `NA`,
         discount = Ds,
         premium = Pr,
         date = DV) %>%
  mutate(id = factor(1:n()),
         type_class = case_when(type_class == 0 ~ "Factory",
                                type_class == 1 ~ "Residence",
                                type_class == 2 ~ "Retail",
                                type_class == 3 ~ "Mixed",
                                type_class == 4 ~ "Build-on Land",
                                type_class == 5 ~ "Office",
                                type_class == 6 ~ "Agricultural Building"
  ),
  type_class = factor(type_class),
  quality = factor(quality,
                   levels = 1:5,
                   labels = c("Poor", "Adequate", "Fair", "Good", "Excellent"),
                   ordered = TRUE),
  state_maintenance = factor(state_maintenance,
                             levels = 1:5,
                             labels = c("Poor", "Adequate", "Fair", "Good", "Excellent"),
                             ordered = TRUE),
  occupancy = factor(occupancy,
                     levels = c(0, 1, 2),
                     labels = c("Unoccupied", "Tenant", "Owner")),
  location = factor(location,
                    levels = c(0, 1, 2),
                    labels = c("Center", "Semi-center", "Suburban")),
  date = as.Date(ymd_hms(date)))
```

Split table by "cluster":
```{r}
auctions_phy <- italy_auctions %>%
  select(id, type_class:state_maintenance)

auctions_pf <- italy_auctions %>%
  select(id, occupancy)

auctions_sef <- italy_auctions %>%
  select(id, location:population)

auctions_amf <- italy_auctions %>%
  select(id, days_on_market:date)
```

Get the column names and labels; these will be used in the documentation:
```{r}
data_names <- data.frame(names = colnames(italy_auctions),
                         label = c("Type of auctioned property; factor with 7 levels (Factory, Residence, Retail, Mixed, Build-on Land, Office, Agricultural Building)",
"Gross Building Area (m2)",
"An ordinal factor that describes the quality of the property, including design work and materials employed (Poor, Adequate, Fair, Good, Excellent)",
"An ordinal factor that describes the state of maintenance of the property, if it is new, recently renovated or damaged over the years (Poor, Adequate, Fair, Good, Excellent)",
"The state of ccupancy of the auctioned property as an indicator of whether the property generates income; factor with 3 levels (Unoccupied, Tenant, Owner)",
"Physical location based on distance from the city center of property, ordinal factor (Center, Semi-center, Suburban)",
"Average per capita income for the city where the auction takes place (euros/year)",
"Variation in the number of local real estate transactions over the observed period, an index that describes the activity of the local real estate market (percentage)",
"Real Estate Activity Index (REAI) is an index that describes the activity and the performance of the local real estate market. This index is calculated as the ratio between delta_ntn and the total stock of real estate units in the area where the property is auctioned",
"Population in the city where the property is auctioned",
"Number of days the property stayed on the auctioned market before being sold; the difference between the date of sale (closing date) and the first auction starting date (opening bid)",
"The number of bidding proceedings before the selling",
"Percentage variation between the first listing value (which equals the market value assessed by the appraisal) and the selling price ",
"Premium paid by the winner bidder to win the auction calculated as the percentage variation between the last listing value and the final selling price",
"Date when the property was assessed by the appraisal, in Year-Month-Day format"
))
```

Write description of variables in roxygen-style:
```{r}
data_names <- data_names %>% mutate(description = glue::glue("#'   \\item{{{names}}}{{{label}}}"))
```

Save tables to file:
```{r}
usethis::use_data(auctions_phy,
                  overwrite = TRUE)

usethis::use_data(auctions_pf,
                  overwrite = TRUE)

usethis::use_data(auctions_sef,
                  overwrite = TRUE)

usethis::use_data(auctions_amf,
                  overwrite = TRUE)
```

