---
title: "Data Preprocessing PhD entrepreneurship in Italy"
output: html_notebook
---

A dataset with information from a survey on PhD graduates and entrepreneurship. Data source:

Muscio, Alessandro, and Laura Ramaciotti. "Dataset from a qualitative survey on Ph. D. entrepreneurship in Italy." Data in brief 18 (2018): 1272-1276. https://doi.org/10.1016/j.dib.2018.03.116

Paper saved as: 1-s2.0-S2352340918303299-main.pdf

In this notebook, the data are pre-processed for packaging in `R`.

Load libraries:
```{r}
library(eurostat)
library(haven)
library(labelled)
library(lubridate)
library(readr)
library(tidyverse)
```

## Original data

Read data file (original is in Stata format):
```{r}
phd_italy <- read_dta("1-s2.0-S2352340918303299-mmc2.zip")
```

Get the column names and labels; these will be used in the documentation:
```{r}
data_names <- purrr::map_df(phd_italy, ~attributes(.x)) %>%
  bind_cols(names = names(phd_italy), .)
```

Write description of variables in roxygen-style:
```{r}
data_names <- data_names %>% mutate(description = glue::glue("#'   \\item{{{names}}}{{{label}}}"))
```

Remove attributes of columns, which will change after converting data to factors:
```{r}
phd_italy <- remove_attributes(phd_italy, c("label", "format.stata"))
```

Preprocess data:
```{r}
phd_italy <- phd_italy %>%
  transmute(id = factor(id),
            q01_annodinascita,
            q02_sesso = factor(q02_sesso, 
                               levels = c(0, 1),
                               labels = c("female", "male")),
            q03_cittadinanzaest = factor(q03_cittadinanzaest,
                                         levels = c(0, 1),
                                         labels = c("Italian", "Foreign")),
            q04_provvive = factor(q04_provvive),
            q05_genitorelaureato = factor(q05_genitorelaureato,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q06_genitoreaccad  = factor(q06_genitoreaccad,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q07_genitoreimpr  = factor(q07_genitoreimpr,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            concludott_cert = factor(q07_genitoreimpr,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q08_phd_clean = factor(q08_phd_clean,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q08a_adj1 = factor(q08a_adj1,
                                          levels = c(0, 1),
                                          labels = c("Yes", "No")),
            q08a_concl_anno = case_when(q08a_adj2 == 1 ~ 2008,
                                        q08a_adj3 == 1 ~ 2009,
                                        q08a_adj4 == 1 ~ 2010,
                                        q08a_adj5 == 1 ~ 2011,
                                        q08a_adj6 == 1 ~ 2012,
                                        q08a_adj7 == 1 ~ 2013,
                                        q08a_adj8 == 1 ~ 2014),
            struttura = factor(struttura),
            code_un = factor(code_un),
            nome_provincia = factor(nome_provincia),
            nome_regione,
            q13a_phdgiudizio,
            q13b_phdgiudizio,
            q13c_phdgiudizio,
            q13d_phdgiudizio,
            q13e_phdgiudizio,
            q13f_phdgiudizio,
            q13g_phdgiudizio,
            q13h_phdgiudizio,
            q13i_phdgiudizio,
            q14_phdfinpriv = factor(q14_phdfinpriv,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q18_phdimprese = factor(q18_phdimprese,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q21_ricbase,
            q29_brevettisn = factor(q29_brevettisn,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q29a_brevettin,
            q30_papernaz2,
            q31_paperint2,
            q371_lavora = factor(q371_lavora,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q372_impresa = factor(q372_impresa,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q372a_impattiva = factor(q372a_impattiva,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q373_impabbandono = factor(q373_impabbandono,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q375_posizuni = factor(q375_posizuni,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q377_posizauto = factor(q377_posizauto,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q3711_impprov = factor(q3711_impprov),
            q3713_impcorso = factor(q3713_impcorso,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            q3715_impaddetti,
            q379_impanni,
            q53a_uniimp,
            q53b_uniimp,
            q53c_uniimp,
            q53d_uniimp,
            q53e_uniimp,
            q53f_uniimp,
            q53g_uniimp,
            q53h_uniimp,
            q53i_uniimp,
            dimensione,
            med_school = factor(med_school,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            polytech = factor(polytech,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            geo = case_when(geo_n == 1 ~ "Northern Italy",
                            geo_c == 1 ~ "Central Italy",
                            geo_s == 1 ~ "Southern Italy"),
            #geo = factor(geo),
            public_uni = factor(public_uni,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            vqr_average,
            utt_sn2006 = factor(utt_sn2006,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            utt_mission_d2006 = factor(utt_mission_d2006,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            disocc2006,
            area = case_when(area1 == 1 ~ "Mathematics and Computer Science",
                             area2 == 1 ~ "Physics",
                             area3 == 1 ~ "Chemistry",
                             area4 == 1 ~ "Geology",
                             area5 == 1 ~ "Biology",
                             area6 == 1 ~ "Medicine",
                             area7 == 1 ~ "Agriculture and Veterinary",
                             area8 == 1 ~ "Civil Engineering and Architecture",
                             area9 == 1 ~ "Engineering",
                             area10 == 1 ~ "Humanities",
                             area11 == 1 ~ "Sociology, Philosophy, and Psychology",
                             area12 == 1 ~ "Law",
                             area13 == 1 ~ "Economics and Statistics",
                             area14 == 1 ~ "Political Sciences"),
            area = factor(area),
            q378_imptipo_clean = fct_explicit_na(q378_imptipo_clean,
                                                 na_level = "(NA)"),
            so_reg2006 = factor(so_reg2006,
                                          levels = c(0, 1),
                                          labels = c("No", "Yes")),
            spinoff0506) %>%
  # Complete missing observations
  mutate(nome_regione = ifelse(nome_provincia == "Milano" & nome_regione == "", "Lombardia", nome_regione),
         nome_regione = ifelse(nome_provincia == "Roma" & nome_regione == "", "Lazio", nome_regione),
         nome_regione = ifelse(nome_provincia == "Siena" & nome_regione == "", "Toscana", nome_regione),
         nome_regione = factor(nome_regione),
         geo = ifelse(nome_regione == "Lazio" & is.na(geo), "Central Italy", geo),
         geo = ifelse(nome_regione == "Lombardia" & is.na(geo), "Northern Italy", geo),
         geo = ifelse(nome_regione == "Toscana" & is.na(geo), "Central Italy", geo),
         geo = factor(geo))
```

Save data:
```{r}
usethis::use_data(phd_italy,
                  overwrite = TRUE)
```

## Regional summary of selected variables

Create a summary of selected variables by region:
```{r}
phd_regional <- phd_italy %>%
  group_by(geo, nome_regione) %>%
  summarize(phd_students = n(),
            active_spinoff = sum(q372a_impattiva == "Yes", 
                                 na.rm = TRUE),
            employees = sum(q3715_impaddetti, 
                            na.rm = TRUE),
            spinoff0506 = sum(spinoff0506,
                              na.rm = TRUE),
            .groups = "drop")
```

Use package {eurostat} to retrieve regional population statistics. Download table and summarize by region for 2005:
```{r}
pop_regional <- get_eurostat(id = "demo_r_d2jan") %>% 
  rename(N2_Code = geo,
         population05 = values) %>%
  filter(time == "2005-01-01", 
         # All ages
         age == "TOTAL",
         # All sexes
         sex == "T",
         # Select Italy
         str_detect(N2_Code, "IT"),
         # Select NUTS-2 regions (with strings of lenght 4)
         str_length(N2_Code) == 4)
```

Read table with NUTS codes:
```{r}
italy_nuts_codes <- read_csv("italy-nuts-codes.csv")
```

Filter NUTS-2 Codes:
```{r}
italy_nuts2_codes <- italy_nuts_codes %>%
  select(NUTS_2, N2_Code) %>%
  distinct(NUTS_2, 
           .keep_all = TRUE)
```

Join to table:
```{r}
pop_regional <- pop_regional %>%
  left_join(italy_nuts2_codes,
            by = "N2_Code") %>%
  transmute(nome_regione = NUTS_2,
            NUTS_ID = N2_Code,
            population05)
```

Join population data to data frame `phd_regional`:
```{r}
phd_italy_regions <- phd_regional %>%
  left_join(pop_regional,
            by = "nome_regione") %>%
  mutate(nome_regione = factor(nome_regione),
         NUTS_ID = factor(NUTS_ID)) %>%
  select(geo,
         nome_regione,
         NUTS_ID,
         phd_students,
         active_spinoff,
         employees,
         spinoff0506,
         population05)
```

Save data:
```{r}
usethis::use_data(phd_italy_regions,
                  overwrite = TRUE)
```

## Geometry

Get geospatial file:
```{r}
italy_nuts2 <- get_eurostat_geospatial(nuts_level = 2, 
                                       year = "2006") %>%
  filter(str_detect(NUTS_ID, "IT")) %>%
  select(-c(id, geo))
```

Plot geometry:
```{r}
ggplot() +
  geom_sf(data = italy_nuts2)
```

Save to file:
```{r}
usethis::use_data(italy_nuts2,
                  overwrite = TRUE)
```

