---
title: "Industrial Readiness Index"
output: html_notebook
---

A dataset with information suitable to measure regional Industry 4.0 (I4.0+) readiness. Data source:

Honti, Gergely, Tímea Czvetkó, and János Abonyi. "Data describing the regional Industry 4.0 readiness index." Data in Brief 33 (2020): 106464. https://doi.org/10.1016/j.dib.2020.106464

Paper saved as: 1-s2.0-S2352340920313469-main.pdf

In this notebook, the data are pre-processed for packaging in `R`.

Load libraries:
```{r}
library(labelled)
library(lubridate)
library(readr)
library(tidyverse)
```

Data saved as: 1-s2.0-S2352340917307138-mmc2.xlsx

Read data file:
```{r}
i40_index_db <- read_delim("I40_indicator_db.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
```

Adjust some names that use problematic syntax:
```{r}
i40_index_varnames <- data.frame(varname = colnames(i40_index_db)) %>%
  mutate(varname = str_replace_all(varname, 
                               "_-_", 
                               "-"),
         varname = str_replace_all(varname,
                               "__",
                               "_"))
```

Replace the names with the clean version:
```{r}
colnames(i40_index_db) <- i40_index_varnames$varname
```


Get the column names and labels; these will be used in the documentation:
```{r}
data_names <- read_delim("I40_indicator_db_column_description.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
```

Copy the column names to data_names:
```{r}
data_names$ColumnName <- i40_index_varnames$varname
```

Write description of variables in roxygen-style:
```{r}
data_names <- data_names %>% mutate(description = glue::glue("#'   \\item{{{ColumnName}}}{{{Description}}}"))
```

Save to file:
```{r}
usethis::use_data(i40_index_db,
                  overwrite = TRUE)
```
